---
title: "Peer-graded Assignment: Statistical Inference Course Project"
author: "Gilbert Toroitich Tarus"
date: "27/01/2021"
output:
  pdf_document:
    df_print: paged
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE,warning=FALSE)
```
### **Overview**
This submission is a report on the investigtion of exponential distribution in R comparing it with the Central Limit Theorem. This project is done using a sample of 40 exponential for thousand simulations.

### **Simulation Example**
**The Exponential Distribution**

In order to simualte 40 observations, we utilize the `rexp()` function in r with arguments `n = 40` and `lamda = .2`. However, to generate 1000 such simulations, we use the `matrix()` function  with dimensions `1000 * 40`

```{r}
nosim <- 1000
n <- 40
lamda <- 0.2
expSimulation <- matrix(rexp(n*nosim,lamda),nosim,n)
theoriticalMean <- 1/lamda
```

Calculate the mean of each sample. We will have 1000 averages.
```{r}
sampledMean <- apply(expSimulation, 1, mean)
```
**1. Show the sample mean and compare it to the theoretical mean of the distribution.**

**Plot**
```{r}
library(ggplot2)
ggplot(data.frame(x <- sampledMean),aes(x = x))+
    geom_histogram(fill = "salmon",bins = 50)+
    geom_vline(xintercept = mean(x),size = 2, color = "black")+
    labs(title = "Histogram of Sample means",x = "Sample Mean")
```

The sample means are are centered arround 5 shown by the vertical line in the plot is equal to theoritical mean `r {theoriticalMean}`.

**2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.**

```{r}
sampledVariance <- apply(expSimulation, 1, var)
ggplot(data.frame(x <- sampledVariance),aes(x = x))+
    geom_histogram(fill = "magenta",bins = 50)+
    geom_vline(xintercept = (1/lamda)^2,size = 2, color = "black")+
    labs(title = "Histogram of Sample Variance",x = "Sample Variance")
```

The distribution of sample variance is centered around the theoritical variance `r {(1/lamda)^2}`

**3. Show that the distribution is approximately normal.**

```{r,fig.height=5,fig.width=12}
library(gridExtra)
set.seed(1234)
x <- rexp(nosim,lamda)
g1 <- ggplot(data.frame(x),aes(x = x))+
    geom_histogram(fill = "salmon",bins = 50)+
    geom_vline(xintercept = mean(x),size = 2, color = "black")+
    labs(title = "Histogram of n = 1000 Samples\nfrom exponential distribution",x = "Sample")
g2 <- ggplot(data.frame(x = sampledMean),aes(x = x))+
    geom_histogram(fill = "salmon",bins = 50)+
    geom_vline(xintercept = mean(x),size = 2, color = "black")+
    labs(title = "Histogram of Sample means for 1000 \nsimulations",x = "Sample Mean")

grid.arrange(g1,g2,ncol = 2)
```

The plot on the left side follows an exponential distribution with mean 5.  
The plot on the right side is bell-shaped. This follows a normal distributin with mean 5 and standard error `r {sd(sampledMean)/sqrt(length(sampledMean))}`

```{r, echo=FALSE,results='hide',fig.show='hide'}

## Example
hist(runif(1000))
mns = NULL
for (i in 1:1000) {
    mns = c(mns,mean(runif(40)))
}

hist(mns)
```